One exact and one random sampling method to assess the reliability of inferences from network enrichment analysis

### 2018-07-31 doc/genesetdp.tex
Prettifying affiliation

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{One exact and one random sampling method to assess the reliability of inferences from network enrichment analysis} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Gustavo S. Jeuken\textsuperscript{1} and
Lukas K\"{a}ll\textsuperscript{1,*}
\\
\bigskip
\textbf{1}Science for Life Laboratory, School of Engineering Sciences in Chemistry, Biotechnology and Health, KTH -- Royal Institute of Technology, Box 1031, 17121 Solna, Sweden
\\
\bigskip

% Use the asterisk to denote corresponding authorship and provide email address in note below.
* lukas.kall@scilifelab.se
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
A prevailing technique to infer function from identifications, from molecular biological high-throughput experiments, is gene set enrichment analysis, where the identifications are compared to predefined sets of related genes often refered to as pathways. As at least some pathways are known to be incomplete in their annotation, algorithmic efforts have been made to complement them with information from functional association networks. While the termonolgy vary in the litterature, we will here refer to such methods as Network Enrichment Analysis (NEA). Traditionally, the significance of inferences from NEA has been assigned using parametric models, that are tuned by fitting parameters by sampling random network perturbations. Here we instead have designed one dynamic programming  algorithm that calculates the score distribution of NEA scores and makes it possible to assign unbiased mid $p$~values to inferences. We also implemented a random sampling method, carrying out the same task. We demonstrate that our method obtains a superior statistical calibration as compared to the popular NEA inference engine, BinoX.

>>>>

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{One exact and one random sampling method to assess the reliability of inferences from network enrichment analysis} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Gustavo S. Jeuken\textsuperscript{1} and
Lukas K\"{a}ll\textsuperscript{1,*}
\\
\bigskip
\textbf{1} Science for Life Laboratory, School of Engineering Sciences in Chemistry, Biotechnology and Health, KTH -- Royal Institute of Technology, Box 1031, 17121 Solna, Sweden
\\
\bigskip

% Use the asterisk to denote corresponding authorship and provide email address in note below.
*lukas.kall@scilifelab.se
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
A prevailing technique to infer function from identifications, from molecular biological high-throughput experiments, is gene set enrichment analysis, where the identifications are compared to predefined sets of related genes often refered to as pathways. As at least some pathways are known to be incomplete in their annotation, algorithmic efforts have been made to complement them with information from functional association networks. While the termonolgy vary in the litterature, we will here refer to such methods as Network Enrichment Analysis (NEA). Traditionally, the significance of inferences from NEA has been assigned using parametric models, that are tuned by fitting parameters by sampling random network perturbations. Here we instead have designed one dynamic programming  algorithm that calculates the score distribution of NEA scores and makes it possible to assign unbiased mid $p$~values to inferences. We also implemented a random sampling method, carrying out the same task. We demonstrate that our method obtains a superior statistical calibration as compared to the popular NEA inference engine, BinoX.
